A pipeline to analyse time-course gene expression data

### 2020-07-27 scripts/manuscript.Rmd
FIX cluster 8 -> cluster 6

```{r topScoreInCluster2, fig.height=6, fig.width=6, fig.cap=paste("Top scoring genes in Cluster",cluster_of_interest)}
# order them by their score in cluster
ord = order(scores_and_labels$scores[genes_in_cluster2,
                     cluster_of_interest])
cluster2Score = genes_in_cluster2[ord]
dataToPlot = rbind(assay(moanin_model[cluster2Score[1:4], ]),
    "Centroid"=kmeans_clusters$centroids[cluster_of_interest, ])
plot_splines_data(moanin_model, data=dataToPlot, 
     colors=ann_colors$Group, smooth=TRUE,simpleY =FALSE,
     mar=c(1.5,2.5,2,0.1))
```

```{r results="hide", echo=FALSE}
# Save the clustering for the GO term and pathway enrichment on the whole set
# of clusters.
kmeans_labels = as.data.frame(labels[!is.na(labels)])
write.table(kmeans_labels,
	    "./results/clustering_labels.txt",
	    sep="\t")
```

@@

Now, let us look more in detail some specific clusters. Cluster 8 seems
particularly interesting: it captures genes with strong differences between
the different influenza treatments and the control, while the control remains
relatively flat.

@@

First, we select the genes of interest, namely those in cluster 8, based on
our goodness-of-fit assignment.

```{r}
cluster_to_plot = 8
genes_to_plot =  names(labels[labels == cluster_to_plot])
```

@@

```{r bootstrapConsensus, fig.width=6, fig.height=4, fig.cap="Consensus matrix of proportion of times samples were in the same cluster over bootstrap samples for k=5 and 20."}
par(mfrow=c(2, 1))
consensus_matrix_stability_5 = consensus_matrix(stability_5,
						scale=FALSE)
aheatmap(consensus_matrix_stability_5[1:1000, 1:1000], Rowv=FALSE,
	 Colv=FALSE,
	 treeheight=0, main="K=5")

@@

## Get gene names of genes in Cluster 8
cluster = 8
labels = unlist(labels)
gene_names = names(labels)
genes = gene_names[labels == cluster]

@@

Then we use the function `find_enriched_pathway` in the `KEGGprofile` package to determine whether any KEGG pathways are enriched in cluster 8, i.e. whether a higher percentage of genes from a single pathway are found in cluster 8 than we would expect by simply proportional assignment of genes to clusters.

@@

**EAP: I changed this to be the min(10, nrow(pathways)) but wondering if there's a problem, since presumably there used to be more than 10 pathways found significant and on my machine there are only 2?**

@@

```{r go_term_enrichment, results="hide", message=FALSE}
# Create logical vector of whether in cluster 8
assignments = labels == cluster

>>>>

```{r topScoreInCluster2, fig.height=6, fig.width=6, fig.cap=paste("Top scoring genes in Cluster",cluster_of_interest)}
# order them by their score in cluster
ord = order(scores_and_labels$scores[genes_in_cluster2,
				     cluster_of_interest])
cluster2Score = genes_in_cluster2[ord]
dataToPlot = rbind(
    assay(moanin_model[cluster2Score[1:4], ]),
    "Centroid"=kmeans_clusters$centroids[cluster_of_interest, ])
plot_splines_data(
    moanin_model, data=dataToPlot, 
    colors=ann_colors$Group, smooth=TRUE, simpleY=FALSE,
    mar=c(1.5, 2.5, 2, 0.1))
```

```{r saving_results, results="hide", echo=FALSE}
# Save the clustering for the GO term and pathway enrichment on the whole set
# of clusters.
kmeans_labels = as.data.frame(labels[!is.na(labels)])
write.table(kmeans_labels,
	    "./results/clustering_labels.txt",
	    sep="\t")
```

@@

Now, let us look more in detail some specific clusters. Cluster 6 seems
particularly interesting: it captures genes with strong differences between
the different influenza treatments and the control, while the control remains
relatively flat.

@@

First, we select the genes of interest, namely those in cluster 6, based on
our goodness-of-fit assignment.

```{r}
cluster_to_plot = 6
genes_to_plot =  names(labels[labels == cluster_to_plot])
```

@@

```{r bootstrapConsensus, fig.width=6, fig.height=4, fig.cap="Consensus matrix of proportion of times samples were in the same cluster over bootstrap samples for k=5 and 20."}
par(mfrow=c(1, 2))
consensus_matrix_stability_5 = consensus_matrix(stability_5,
						scale=FALSE)
aheatmap(consensus_matrix_stability_5[1:1000, 1:1000], Rowv=FALSE,
	 Colv=FALSE,
	 treeheight=0, main="K=5")

@@

## Get gene names of genes in Cluster 8
cluster = 6
gene_names = names(labels)
genes = gene_names[labels == cluster]

@@

Then we use the function `find_enriched_pathway` in the `KEGGprofile` package
to determine whether any KEGG pathways are enriched in cluster 6, i.e. whether
a higher percentage of genes from a single pathway are found in cluster 6 than
we would expect by simply proportional assignment of genes to clusters.

@@

@@

```{r go_term_enrichment, results="hide", message=FALSE}
# Create logical vector of whether in cluster 6
assignments = labels == cluster
